### 2022-09-28 response_to_reviewers.Rmd
fixed most of #25

---
title: "Response to Reviewers"
date: "2022-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Major comments from Reviewer 1

Comments to the Author(s)
Please see attached file for comments.

## Major comments from Reviewer 2

> While Sections 2 and 3 are well done and complete, Sections 4 and 5 are weak and seem incomplete and unpolished.

>Likewise the Opportunities section should be expanded to include the proliferation of Sports Analytics conferences.

> My largest concern is that a cursory discussion of player contribution or player ratings is given and this seems a major oversight of one of the “big ideas” in sports analytics given the focus on this topic in many sports.  See, for example, Mark Glickman’s recent presentation on this at the 2022 Joint Statistical Meetings.  I don’t think that Hvattum (2019) is sufficient to cover the area of player contribution.

## Minor comments from Reviewer 2

> Pages 3,4, 5: The use of historical tables and data is powerful.  However, I worry about the quality of the images particularly Figure 2.

We share this concern. We are also uncertain about any licensing restrictions. We have left the images as is for now, and will wait direction from Wiley.

> Page 6, line 8:  “the effect may be overstated” is somewhat ambiguous.  Is it the

XX Quang or Greg to respond

> Page 6, line 53: Is the “!” warranted here for conditioning on a 9-dimensional vector?

We have replaced the exclamation point with a period.

> Page 8, line 53: There is also work on expected possession in ice hockey, see Stokes et al. https://hockey-graphs.com/2021/07/06/bayesian-space-time-models-for-expected-possession-added-value-part-1-of-2/

We have added a reference to this work, but are also cognizant that Wiley policy may not permit a citation to this unpublished work:

>> "preprint articles should not be cited in WIREs manuscripts as review articles should discuss and draw conclusions only from peer-reviewed research. Remember that original research/unpublished work should also not be included as it has not yet been peer-reviewed and could put the work in jeopardy of getting published in the primary press."

> 9, 17: Is Player Win Average relevant to win probability?  This is more of a player contribution metric than a win probability metric.  I would rather see a brief exposition of Tango et al (2007) here.

Given that Albert (2015) previously published an article on this topic in this same journal series, we did not include a treatment of player contribution metrics in our paper. Our opinion is that this is redundant given Albert's article.

> 9, 47: Gridiron football is a lovely term but for this audience, I think it needs to be clarified.

We like the term too, but have changed this to "American football."

> 10, 52: “In this paper”, the “this” seems ambiguous.

We have replaced this with "Deshpande & Jensen"

> 11, 36: Table 3, why order this table alphabetically, why not by ability level?

We have ordered the rows of Table 3 by descending order of ability.

> 11, 50: Given how much fivethirtyeight.com uses Elo ratings for all kinds of sports, it is probably worth mentioning their work as examples.

We have added the sentence: "Elo ratings are used extensively for rankings of teams in numerous sports by the data journalists at FiveThirtyEight.com." But again, we acknowledge that Wiley policy may not allow such a reference.

> 12, 1: As noted above, Section 4 feels much weaker than Sections 2 and 3 which are strong.  Perhaps a look at Pythagorean methods as a historical introduction to this topic.

XX We could consider this.

> 12, 23: This is not much of a case study.  Perhaps some additional information on the data available, ways it could be used to calculate game state expected valued, win probability, player ratings….

XX Yeah

> 12, 37: Theming *teamcolors* is a great *R* package, not sure it is part of the “big ideas” in sports analytics.  A single line in the first paragraph of this section would seem to be better.

We have added examples from the `SportyR` package, which draws playing surfaces, to help flesh out this subsection.

> 13, 35: Add a citation for Big Data Cup, https://www.stathletes.com/big-data-cup/, for SABR Diamond Dollars Case Competition, https://sabr.org/analytics/case, Big Data Derby https://www.kaggle.com/competitions/big-data-derby-2022, and maybe more.

Thank you for these suggestions. We have added them.

> 13, 43-53: don’t think there is a need for a bulleted list here.