### 2023-01-03 response_to_reviewers_R1.Rmd
add chess figures; #44

---
title: "Response to Reviewers"
date: "2023-01-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Major comments from Reviewer 1

## Major comments from Reviewer 2

Overall Summary:

This revised version is a significant improvement over the previous version and the reviewers have addressed my major concerns with the paper with one exception.

Suggested Changes:

The authors have now made clearer the limitation in scope of this manuscript related to player evaluation. Perhaps, a statement regarding this can be added to the abstract.

I think that parts of the first paragraph of Section 5 regarding biometric data is a bit out of place. This could be moved to Section 7 (Opportunities). Likewise, Section 5 could be retitled something like‘Betting Data and Sports Analytics” or “Betting Data Analysis”.

## Minor Changes from Reviewer 2

- Page 1, line 47 “And, of course, in baseball” add commas before and - after “of course”

> Done.

- Page 2, line 5, “games share” instead of “game shares”

> Fixed.

- Page 2 line 6, “anthropologically’ seems odd here. Maybe “historically” instead

> Fixed.

- Page 8, Macdonald (2012?) used expected possession value for ice hockey to evaluate players.

> We have a added a citation to this work, but note that whereas Macdonald is using expected *goals*, he did not have access to player tracking data necessary to compute an EPV metric.

- Section 6.1, line 53, Figure ??

> This is an artifact of the attempt to track changes and does not appear in the unannotated manuscript.

- Section 6.1, line 56, contextual is used twice in that sentence, perhaps “helping the reader visualize” or “helping the reader understand”

> We have changed this to "understand".

- Table 3, the third column of this table is off by a line with the entries in the first two columns.

> This is an artifact of the attempt to track changes and does not appear in the unannotated manuscript.

- Section 6.2, line 52, Figure 6 should be Figure 7; likewise the figure number in the next sentence

> This is a downstream consequence of the previous error with Figure ??

- Figure 6, in my version the images are stacked rather than left/right

> This is an artifact of the attempt to track changes and does not appear in the unannotated manuscript.

- Section 6.2, line 46 This sentence that starts “Consider, for example, “ seems out of place with the rest of the tone of this article

> TODO

- Section 7, line 7 or 8, it might be worth mentioning the growth of small conferences like UCSAS that have sprung up, something like “As the field of sports analytics has grown, there has been a proliferation of sports specific and regional sports analytics conferences.”

> We have added a similar sentence to this effect.