### 2023-01-03 response_to_reviewers_R1.Rmd
address R1; #43

---
title: "Response to Reviewers"
date: "2023-01-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Major comments from Reviewer 1

Overall, the paper is now much more comprehensive, complete, and clear; I think it
will serve as a very useful introduction to the field of sports analytics. I have only a few
suggestions and points of confusion to share, which I hope the authors will examine before
submitting the final version. They are all fairly minor.

- If there is still space allowed in the abstract, I think you could consider explicitly
mentioning the three big ideas (expected value of a game state, win probability, and
measures of team strength). This would give readers a better idea of what is to come
and also allow people to find your article more easily if they are interested in one of
those specific topics.
- In Section 2.1.2, I am a little confused as to what Carroll et. al. did to extend the
previous model of Carter and Machol, since both efforts seem to involve some sort of
linearity assumption.
- Also in Section 2.1.2, it might be worthwhile to briefly define absorbing states, as I for
one didn’t know the definition before looking it up (and as it turns out the concept is
quite simple)
- In Section 2.1.3, you make the comment “they estimate that a better strategy would
be worth about 0.4 wins per season on average, a substantial amount in a season with
only 16 games.” This makes sense, but previously in Section 2.1.2 you mentioned
a potential difference of 3.8 wins per season, yet didn’t comment on the significant
magnitude of that effect. I’m not advocating for a particular change here, but the
authors could consider adding a small comment about the size of the effect Pelechrinis
et. al. estimated.
- At the end of Section 2.1.3, you write “Lopez (2020) also concludes that coaches are
too conservative on 4th down decision making, but the effect may be overstated due
to the way in which to go yardage has been recorded in the past.” It’s unclear to me
whether the overstated effect is the one Lopez estimates in his paper, or if you are
referring to the estimated effects past studies which used less granular data.
- In Section 2.3, you write “NHL coaches do not always employ the optimal strategies”,
but it isn’t clear what those strategies are in reference to the previous discussion. Is
pulling the goalie optimal?
- In the second formula of Section 4.1, the right hand side of the equation has a 2, but
it should be an α.

> Thank you. This is fixed.

- I think Section 5.2 would benefit from a short introductory paragraph, perhaps which
introduces different types of sports betting, thus leading into the betting pools discus-
sion of Section 5.2.1.
- I notice you have several hyperlinks throughout–I’m not sure what the Wiley standards
are, but it might be helpful to distinguish links with a different color.

> Indeed. We await Wiley's instructions when we reach the typesetting phase.

## Minor comments from Reviewer 1

- Page 5, lines 33-34: ‘might reasonably’ is used twice in the same sentence

> We have changed the first "might reasonably" to "could"

- Page 6, line 23: ‘topics’ instead of ‘topic’

> Fixed.

- Page 6, line 38: ‘go for it’ instead of ‘got for it’

> Fixed.

- Page 6, lines 45-46: ‘coaches’ instead of ‘coached’ and ‘decision making’ instead of
‘decisions making’

> Fixed.

- Page 7, line 7: ‘kickoffs’ instead of ‘kickoff’

> Fixed.

- Page 8, line 51: Might be better to say “Earlier in Section 2” as opposed to just “In
Section 2”

> Fixed.

- Page 11, line 56 and Page 12, line 5: ‘win-loss’ instead of ‘won-loss’

> Fixed.

- Page 13, line 53: This is probably a matter of personal preference but “rating less than
2300” sounds more natural than “rating fewer than 2300”

> Fixed.

- Page 14, line 35: ‘use’ instead of ‘uses’

> Fixed.

- Page 17, line 27: I think the question in parentheses should be placed before the period
of the previous sentence

> We moved this.

- Page 18, line 40: ‘Lindsey’ instead of ‘Lindsay’

> Fixed.

- Page 20, lines 13-15: Not sure why there are brackets around this sentence.

> We removed the brackets.

- Page 20, line 41: ‘Its’ not ‘It’s’

> Fixed.

## Major comments from Reviewer 2

Overall Summary:

This revised version is a significant improvement over the previous version and the reviewers have addressed my major concerns with the paper with one exception.

Suggested Changes:

The authors have now made clearer the limitation in scope of this manuscript related to player evaluation. Perhaps, a statement regarding this can be added to the abstract.

I think that parts of the first paragraph of Section 5 regarding biometric data is a bit out of place. This could be moved to Section 7 (Opportunities). Likewise, Section 5 could be retitled something like‘Betting Data and Sports Analytics” or “Betting Data Analysis”.

## Minor Changes from Reviewer 2

- Page 1, line 47 “And, of course, in baseball” add commas before and - after “of course”

> Done.

- Page 2, line 5, “games share” instead of “game shares”

> Fixed.

- Page 2 line 6, “anthropologically’ seems odd here. Maybe “historically” instead

> Fixed.

- Page 8, Macdonald (2012?) used expected possession value for ice hockey to evaluate players.

> We have a added a citation to this work, but note that whereas Macdonald is using expected *goals*, he did not have access to player tracking data necessary to compute an EPV metric.

- Section 6.1, line 53, Figure ??

> This is an artifact of the attempt to track changes and does not appear in the unannotated manuscript.

- Section 6.1, line 56, contextual is used twice in that sentence, perhaps “helping the reader visualize” or “helping the reader understand”

> We have changed this to "understand".

- Table 3, the third column of this table is off by a line with the entries in the first two columns.

> This is an artifact of the attempt to track changes and does not appear in the unannotated manuscript.

- Section 6.2, line 52, Figure 6 should be Figure 7; likewise the figure number in the next sentence

> This is a downstream consequence of the previous error with Figure ??

- Figure 6, in my version the images are stacked rather than left/right

> This is an artifact of the attempt to track changes and does not appear in the unannotated manuscript.

- Section 6.2, line 46 This sentence that starts “Consider, for example, “ seems out of place with the rest of the tone of this article

> TODO

- Section 7, line 7 or 8, it might be worth mentioning the growth of small conferences like UCSAS that have sprung up, something like “As the field of sports analytics has grown, there has been a proliferation of sports specific and regional sports analytics conferences.”

> We have added a similar sentence to this effect.